---
title: "Class 17: Investigating Pertussis Resurgence"
author: "Soomin Park"
format: pdf
---

Pertussis, or whooping cough, is a highly contagious lung infection caused by a bacteria *B. pertussis*

The CDC tracks reported cases in the U.S. since the 1920s. 

## 1. Investigating pertussis cases by year

> Q1. With the help of the R “addin” package datapasta assign the CDC pertussis case number data to a data frame called cdc and use ggplot to make a plot of cases numbers over time.

```{r}
cdc <- data.frame(
                                 Year = c(1922L,1923L,1924L,1925L,
                                          1926L,1927L,1928L,1929L,1930L,1931L,
                                          1932L,1933L,1934L,1935L,1936L,
                                          1937L,1938L,1939L,1940L,1941L,1942L,
                                          1943L,1944L,1945L,1946L,1947L,
                                          1948L,1949L,1950L,1951L,1952L,
                                          1953L,1954L,1955L,1956L,1957L,1958L,
                                          1959L,1960L,1961L,1962L,1963L,
                                          1964L,1965L,1966L,1967L,1968L,1969L,
                                          1970L,1971L,1972L,1973L,1974L,
                                          1975L,1976L,1977L,1978L,1979L,1980L,
                                          1981L,1982L,1983L,1984L,1985L,
                                          1986L,1987L,1988L,1989L,1990L,
                                          1991L,1992L,1993L,1994L,1995L,1996L,
                                          1997L,1998L,1999L,2000L,2001L,
                                          2002L,2003L,2004L,2005L,2006L,2007L,
                                          2008L,2009L,2010L,2011L,2012L,
                                          2013L,2014L,2015L,2016L,2017L,2018L,
                                          2019L,2020L,2021L),
         Cases = c(107473,164191,165418,152003,
                                          202210,181411,161799,197371,
                                          166914,172559,215343,179135,265269,
                                          180518,147237,214652,227319,103188,
                                          183866,222202,191383,191890,109873,
                                          133792,109860,156517,74715,69479,
                                          120718,68687,45030,37129,60886,
                                          62786,31732,28295,32148,40005,
                                          14809,11468,17749,17135,13005,6799,
                                          7717,9718,4810,3285,4249,3036,
                                          3287,1759,2402,1738,1010,2177,2063,
                                          1623,1730,1248,1895,2463,2276,
                                          3589,4195,2823,3450,4157,4570,
                                          2719,4083,6586,4617,5137,7796,6564,
                                          7405,7298,7867,7580,9771,11647,
                                          25827,25616,15632,10454,13278,
                                          16858,27550,18719,48277,28639,32971,
                                          20762,17972,18975,15609,18617,
                                          6124,2116)
       )
```

We can now plot the number of reported cases 

```{r}
library(ggplot2)

ggplot(cdc) +
  aes(Year, Cases) +
  geom_point() +
  geom_line() +
  labs(x = "Year", y = "Number of Cases", title = "Pertussis Cases by Year (1922-2021)")

```

## 2. A tale of two vaccines (wP & aP)

The first big "whole-cell" vaccine program started in 1942.

> Q2. Using the ggplot geom_vline() function add lines to your previous plot for the 1946 introduction of the wP vaccine and the 1996 switch to aP vaccine (see example in the hint below). What do you notice?

```{r}
ggplot(cdc) +
  aes(Year, Cases) +
  geom_point() +
  geom_line() +
  labs(x = "Year", y = "Number of Cases", title = "Pertussis Cases by Year (1922-2021)") + 
  geom_vline(xintercept=1942, color = "blue") +
  geom_vline(xintercept=1980, color = "gray", linetype=2) +
  geom_vline(xintercept=1995, color = "red")
```

> Q3. Describe what happened after the introduction of the aP vaccine? Do you have a possible explanation for the observed trend?

Something big is happening with pertussis cases and big outbreaks are once again a major public health concern. 

One of the main hypothesis for the increasing case numbers is waning vaccine efficiency with the newer aP vaccine. 

## 3. Exploring CMI-PB data

Enter the CMI-PB project, which is studying this problem on large scale. Let's see what data they have. 

Their data is available in JSON format ("key:value" pair style).
We will use the "jsonlite" package to read their data. 

```{r}
library(jsonlite)

subject <- read_json("https://www.cmi-pb.org/api/subject", simplifyVector = TRUE) 

head(subject, 3)
```
> Q4. How many aP and wP infancy vaccinated subjects are in the dataset?

```{r}
table(subject$infancy_vac)
```

> Q5. How many Male and Female subjects/patients are in the dataset?

```{r}
table(subject$biological_sex)
```

```{r}
table(subject$race)
```

> Q6. What is the breakdown of race and biological sex (e.g. number of Asian females, White males etc…)?

```{r}
table(subject$race, subject$biological_sex)
```

Now let's read some more database tables from CMI-PB

```{r}
specimen <- read_json("http://cmi-pb.org/api/specimen", simplifyVector = TRUE)

head(specimen)
```

I want to "join" (a.k.a "merge"/link/etc.) the `subject` and `speciment` tables together.
I will use the **dplyr** package for this. 

> Q9. Complete the code to join specimen and subject tables to make a new merged data frame containing all specimen records along with their associated subject details:

```{r}
library(dplyr)

meta <- inner_join(subject, specimen)
dim(meta)
head(meta)
```

```{r}
ab <- read_json("https://www.cmi-pb.org/api/ab_titer", simplifyVector = TRUE) 

head(ab)
```

Now I can join `meta` that we made above which contains all info about the subjects and speciemens with this `ab` data.

> Q10. Now using the same procedure join meta with titer data so we can further analyze this data in terms of time of visit aP/wP, male/female etc.

```{r}
abdata <- inner_join(meta, ab)
dim(abdata)
```

```{r}
head(abdata)
```

> Q11. How many specimens (i.e. entries in abdata) do we have for each isotype?

```{r}
table(abdata$isotype)
```

> Q12. What do you notice about the number of visit 8 specimens compared to other visits?

```{r}
table(abdata$visit)
```

There are way less visit 8 specimens because the project is still ongoing and we have not got that data for all individuals yet. 


## 4. Examine IgG1 Ab titer levels

We will use the `filter` function from dplyr to focus on just IgG1 isotype and visits 1 to 7 (i.e. exclude visit 8 as there are not many specimens there yet). 

```{r}
ig1 <- abdata %>% filter(isotype == "IgG1", visit!=8)
head(ig1)
```

Box plot of antigen levels over timel

> Q13. Complete the following code to make a summary boxplot of Ab titer levels for all antigens:

```{r}
library(ggplot2)

ggplot(ig1) +
  aes(MFI, antigen) +
  geom_boxplot()
```

and facet by visit:

```{r}
ggplot(ig1) +
  aes(MFI, antigen) +
  geom_boxplot() + 
  facet_wrap(vars(visit), nrow=2)
```

> Q14. What antigens show differences in the level of IgG1 antibody titers recognizing them over time? Why these and not others?

FIM2/3 changes. This is "Fimbrial protein" that makes the bacteria pilus and is involved in cell adhesion. 

PT Pertussis toxin

FHA is Filamentous hemagglutinin surface associated adherence protein of bacteria pertussis, which is a components of some new acellular pertussis vaccines. 

We can attempt to examine differences between wP and aP here by setting color and/or facet values of the plot to include infancy_vac status 

```{r}
ggplot(ig1) +
  aes(MFI, antigen, col=infancy_vac ) +
  geom_boxplot(show.legend = FALSE) + 
  facet_wrap(vars(visit), nrow=2) +
  theme_bw()
```

Another version of this plot adding infancy_vac to the faceting:

```{r}
ggplot(ig1) +
  aes(MFI, antigen, col=infancy_vac ) +
  geom_boxplot(show.legend = FALSE) + 
  facet_wrap(vars(infancy_vac, visit), nrow=2)
```

> Q15. Filter to pull out only two specific antigens for analysis and create a boxplot for each. You can chose any you like. Below I picked a “control” antigen (“Measles”, that is not in our vaccines) and a clear antigen of interest (“FIM2/3”, extra-cellular fimbriae proteins from B. pertussis that participate in substrate attachment).

```{r}
filter(ig1, antigen=="Measles") %>%
  ggplot() +
  aes(MFI, col=infancy_vac) +
  geom_boxplot(show.legend = FALSE) +
  facet_wrap(vars(visit)) +
  theme_bw()
```

> and the same for antigen=="FIM2/3"

```{r}
filter(ig1, antigen=="FIM2/3") %>%
  ggplot() +
  aes(MFI, col=infancy_vac) +
  geom_boxplot(show.legend = FALSE) +
  facet_wrap(vars(visit)) +
  theme_bw()
```

