---
output:
  pdf_document: default
  html_document: default
---
## Section 4: Population Scale Analysis 

One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 

So, you processed about ~230 samples and did the normalization on a genome level. Now, you want to find whether there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3 expression.

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```

The total sample size of the dataset is 462.

```{r}
table(expr$geno)
```

A/A genotype has the sample size of 108.
A/G genotype has the sample size of 233.
G/G genotype has the sample size of 121.


```{r}
summary(expr$exp)
```

```{r}
aggregate(exp ~ geno, data = expr, FUN = median)
```

Median expression level of the A/A genotype is 31.25.
Median expression level of the A/G genotype is 25.06
Median expression level of the G/G genotype is 20.07

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

```{r}
library(ggplot2)
```


```{r}
ggplot(expr) + aes(geno, exp, col=geno, fill=geno) +
  geom_boxplot(notch=TRUE)
```

The boxplot shows that expression is a lot higher with A/A genotype compared to G/G genotype. Thus, the SNP affects the expression of ORMDL3.

