---
title: "class 10"
author: "Soomin Park"
format: pdf
---

# Comparative analysis of ADK

ADK (Adenelate Kinase) is an important drug target and we would love to know how it works - i.e. molecular mechanism.

There has been lots of work done on this protein due to its importance including lots of crystal structures.

> Q10. Which of the packages above is found only on BioConductor and not CRAN? 

msa

> Q11. Which of the above packages is not found on BioConductor or CRAN?: 

bio3d-view

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? 

TRUE

```{r}
library(bio3d)

aa <- get.seq("1ake_A")
aa
```

> Q13. How many amino acids are in this sequence, i.e. how long is this sequence?

214


We can now run BLAST with this sequence 

```{r}
#b <- blast.pdb(aa)
```

Plot a summary of search results

```{r}
#hits <- plot(b)
```

Let's see what is in our `hits' object

```{r}
#hits$pdb.id

hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```


Now we can download all these PDB structure files

```{r}
#Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```


Now I want to align and superpose these structures which are all over the place. 

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

Let's have a look at our `pdbs` object:

```{r}
pdbs
```

Now we have our aligned and superposed structures we can perform all sorts of analysis on them. Let's do PCA...

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

Results of PCA on Adenlate Kinasee X-ray structures. Each dot represents one PDB structure.

We can cluster the structures by RMSD (or any other method)

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col=grps)
```

We can make a movie - also called a trajectory of the major differences (i.e. structural displacvements) of ADK

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```



















